---
title: "Generate Multiple Site Reports"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document explores how to generate multiple data collection site reports.

# Preparatory steps

We need to determine what counties in what states are plausible recruitment targets for each site.
We'll use a stored `.csv` for the initial testing.

```{r}
site_data <- readr::read_csv("analysis/csv/city-state-county.csv")
```

Some sites have multiple counties, but we'll pick the easier set of those with a single county.

```{r}
single_county_sites <- site_data %>%
  dplyr::filter(multi == "no")
```

# Generate single report

We will make a few single site reports to test the functionality.

```{r}
site_1 <- single_county_sites[1,]

site_1
```

## Site `r site_1$Site.code`

```{r}
rmarkdown::render("PLAY-recruiting-site-report.Rmd", params = list(state = site_1$State, counties=site_1$County), output_file = paste0("PLAY-site-report-", site_1$Site.code, ".html"))
```

This demonstrates that we can script the generation of multiple site reports.
