---
title: "Data collection site demographics"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output:
  github_document:
  html_document:
    toc: true
    toc_levels: 2
    toc_floating: true
    code_folding: hide
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install packman for subsequent package management
if (!("pacman" %in% installed.packages()[,])) {
    install.packages("pacman")
}

# Load packages
pacman::p_load(ggplot2, dplyr, stringr, choroplethr, choroplethrMaps)

# Paths
csv.dir <- "analysis/csv/"
```

## Background

This report provides preliminary analysis of the demographic characteristics of the sites under consideration to collect data for the PLAY project. This report focuses on data at the county level. It uses the [`choroplethr`](https://cran.r-project.org/web/packages/choroplethr/index.html) package.

## Generate data file(s)

Load county names, and add county FIPS from `choroplethr` package's `county.regions` dataset. With this, I can extract demographic data using the `get_county_demographics()` function.

```{r generate-data-files}
counties <- read.csv(paste0(csv.dir, "city-state-county.csv"), stringsAsFactors = FALSE)

# convert counties so can pull FIPS codes
counties$County <- tolower(counties$County)

# Load county data from choroplethr
# Could also use acs package to get updated info.
data("county.regions")

counties <- left_join(counties, county.regions,
                         by = c("County" = "county.name",
                                "State" = "state.abb"))

demo <- get_county_demographics(endyear=2010, span=5)

county.demo <- left_join(counties, demo)
str(county.demo)

# Function to add initial capital
Init.cap <- function(s) {
  stopifnot(is.character(s))
  paste0(toupper(str_sub(s,1,1)), str_sub(s,2,str_length(s)))
}

# Recapitalize county
county.demo$County <- Init.cap(county.demo$County)
```

## Tabular summary

```{r tabular-summary}
county.demo %>%
  arrange(US.Region, State, City, County) %>%
  select(US.Region, City, State, County, total_population, percent_white,
         percent_black, percent_asian, percent_hispanic) %>%
  knitr::kable()
```

## Population by region

```{r pop-by-region-boxplot}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = total_population) +
  geom_violin() +
  geom_point() +
  ggtitle("Population by region")
```

## Race & ethnicity

```{r black}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_black) +
  geom_violin() +
  geom_point() +
  ggtitle("African American population by region") +
  ylab("% African American")
```
```{r hispanic}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_hispanic) +
  geom_violin() +
  geom_point() +
  ggtitle("Hispanic population by region") +
  ylab("% Hispanic")
```

```{r asian}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_asian) +
  geom_violin() +
  geom_point() +
  ggtitle("Asian population by region") +
  ylab("% Asian")
```

```{r black-hispanic-region}
county.demo %>%
  ggplot() +
  aes(x = percent_black, y = percent_hispanic, 
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Percent African American vs. Hispanic") +
  xlab("% African American") +
  ylab("% Hispanic")
```

## Economic indicators

```{r income-rent-region}
county.demo %>%
  ggplot() +
  aes(x = per_capita_income, y = median_rent,
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Median rent by per capita income")
```

## Resources

### R Session

```{r session-info}
sessionInfo()
```
