---
title: "Data collection site demographics"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output:
  github_document:
  html_document:
    toc: true
    toc_levels: 2
    toc_floating: true
    code_folding: hide
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      error = FALSE,
                      message = FALSE,
                      fig.path = "img/")

# Install packman for subsequent package management
if (!("pacman" %in% installed.packages()[,])) {
    install.packages("pacman")
}

# Load packages
pacman::p_load(ggplot2, dplyr, stringr, choroplethr, choroplethrMaps)

# Paths
csv.dir <- "analysis/csv/"

# Helper functions
source("analysis/R/Cap.all.R")
```

## Background

This report provides preliminary analysis of the demographic characteristics of the sites under consideration to collect data for the PLAY project. This report focuses on data at the county level. It uses the [`choroplethr`](https://cran.r-project.org/web/packages/choroplethr/index.html) package.

## Generate data file(s)

Load county names, and add county FIPS from `choroplethr` package's `county.regions` dataset. With this, I can extract demographic data using the `get_county_demographics()` function.

```{r generate-data-files}
counties <- read.csv(paste0(csv.dir, "city-state-county.csv"), stringsAsFactors = FALSE)

# Load county data from choroplethr
# Could also use acs package to get updated info.
data("county.regions")

counties <- left_join(counties, county.regions)

demo <- get_county_demographics(endyear=2013, span=5)

county.demo <- left_join(counties, demo)

# Recapitalize county
county.demo$County <- unlist(lapply(county.demo$County, Cap.all))
```

## Tabular summary

```{r tabular-summary}
# Select collecting sites only
county.demo %>%
  filter(Collecting == "Collecting") ->
  county.demo

county.demo %>%
  arrange(US.Region, Site.code, State, City, County) %>%
  select(US.Region, Site.code, City, State, County, total_population, percent_white,
         percent_black, percent_asian, percent_hispanic) %>%
  knitr::kable()
```

### Summaries across sites

```{r race-ethnicity-across-sites}
county.demo %>%
  arrange(US.Region, Site.code, State, City, County) %>%
  select(US.Region, Site.code, City, State, County, total_population, percent_white,
         percent_black, percent_asian, percent_hispanic) %>%
  group_by(US.Region) %>%
  summarise(med_black = median(percent_black),
            min_black = min(percent_black),
            max_black = max(percent_black),
            med_hisp = median(percent_hispanic),
            min_hisp = min(percent_hispanic),
            max_hisp = max(percent_hispanic))%>%
  knitr::kable()
```

## Population by region

```{r pop-by-region-boxplot}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = total_population) +
  geom_violin() +
  geom_point() +
  ggtitle("Population by region")
```

## Race, ethnicity, & age

```{r black}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_black) +
  geom_violin() +
  geom_point() +
  ggtitle("African American population by region") +
  ylab("% African American")
```
```{r hispanic}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_hispanic) +
  geom_violin() +
  geom_point() +
  ggtitle("Hispanic population by region") +
  ylab("% Hispanic")
```

```{r asian}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = percent_asian) +
  geom_violin() +
  geom_point() +
  ggtitle("Asian population by region") +
  ylab("% Asian")
```

```{r black-hispanic-region}
county.demo %>%
  ggplot() +
  aes(x = percent_black, y = percent_hispanic, 
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Percent African American vs. Hispanic") +
  xlab("% African American") +
  ylab("% Hispanic")
```

```{r black-asian-region}
county.demo %>%
  ggplot() +
  aes(x = percent_black, y = percent_asian, 
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Percent African American vs. Asian") +
  xlab("% African American") +
  ylab("% Asian")
```

```{r hispanic-asian-region}
county.demo %>%
  ggplot() +
  aes(x = percent_hispanic, y = percent_asian, 
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Percent Hispanic vs. Asian") +
  xlab("% Hispanic") +
  ylab("% Asian")
```

```{r age}
county.demo %>%
  ggplot() +
  aes(x = US.Region, y = median_age) +
  geom_violin() +
  geom_point() +
  ggtitle("Median age by region") +
  ylab("Age (yrs)")
```

## Economic indicators

```{r econ-tabular-summary}
county.demo %>%
  arrange(US.Region, Site.code, State, City, County) %>%
  select(US.Region, Site.code, City, State, County, total_population,
         per_capita_income, median_rent) %>%
  knitr::kable()
```

### Summary across counties

This plots the median (min & max) of the per capita income and the median (min & max) of median rent.

```{r econ-per-capita-income-table}
county.demo %>%
  arrange(US.Region, Site.code, State, City, County) %>%
  select(US.Region, Site.code, City, State, County, total_population,
         per_capita_income, median_rent) %>%
  group_by(US.Region) %>%
  summarize(med_per_cap_inc = median(per_capita_income),
            min_per_cap_inc = min(per_capita_income),
            max_per_cap_inc = max(per_capita_income),
            med_med_rent = median(median_rent),
            min_med_rent = min(median_rent),
            max_med_rent = max(median_rent)) %>%
  knitr::kable()

```


```{r income-rent-region}
county.demo %>%
  ggplot() +
  aes(x = per_capita_income, y = median_rent,
      color = US.Region,
      size = total_population) +
  geom_point() +
  ggtitle("Median rent by per capita income")
```

## Further exploration of the ACS

The `acs` package manual can be found here: <https://cran.r-project.org/web/packages/acs/acs.pdf>. A useful guide to the ACS can be found here: <https://www.census.gov/content/dam/Census/library/publications/2008/acs/ACSGeneralHandbook.pdf>. And, an especially useful guide by the `acs` package author can be found here: <http://dusp.mit.edu/sites/dusp.mit.edu/files/attachments/publications/working_with_acs_R_v_2.0.pdf>

### Generate geography for sites

As a first attempt, create vectors of numeric state FIPS and county FIPS codes.

```{r generate-geo}
state.fips <- as.numeric(county.demo$state.fips.character)
county.fips <- as.numeric(substr(county.demo$county.fips.character,3,5))
play.geo <- geo.make(state = state.fips, county = county.fips)
```

This works, but after some further exploration, I think it may be better to generate site-specific geographies.

```{r generate-site-geo}
Make.county.geo <- function(i, df) {
  geo.make(state = as.numeric(df$state.fips.character[i]),
              county =
             as.numeric(substr(county.demo$county.fips.character[i],4, 6)))
}

cty <- 1
# Generate name for county-level geography
geo.name <- paste0(county.demo$Site.code[cty], "_", county.demo$county.name[cty], "_", county.demo$State[cty])

# Create geography and assign to generated name
assign(geo.name, Make.county.geo(cty, county.demo))
```

### Education data

```{r ed-attain-all-counties}
ed.attain <- acs.lookup(table.name="Educational Attainment for the Population 25 Years and Over", endyear=2015)

# Variables 1:25 seem to contain the relevant info
play.ed <- acs.fetch(geography = play.geo, endyear = 2015, variable = ed.attain[1:25],
                     col.names = c("Total",
                                   "None",
                                   "<K",
                                   "K",
                                   "1st",
                                   "2nd",
                                   "3rd",
                                   "4th",
                                   "5th",
                                   "6th",
                                   "7th",
                                   "8th",
                                   "9th",
                                   "10th",
                                   "11th",
                                   "12th",
                                   "HS",
                                   "GED",
                                   "Coll <1yr",
                                   "Coll >1yr",
                                   "AA",
                                   "BA",
                                   "MA",
                                   "Prof",
                                   "Ph.D"))

# Let's collapse for row/county = 1

(lt.hs <- sum(play.ed[1,2:16]))
(hs.grad <- sum(play.ed[1,17:18]))
(some.coll <- sum(play.ed[1,19:21]))
(BA <- sum(play.ed[1,22]))
(MA.plus <- sum(play.ed[1,23:25]))
```

```{r educational-attainment-redux}
# Table B15003
# ed.table.B15003 <- acs.fetch(geography = play.geo, endyear = 2015, table.number = "B15003")
# 
# ed.table.B15003
```

```{r ed-attain-tables}
# Columns 2:16 are grades < HS diploma

lt.hs <- function(i) sum(play.ed[i,2:16])
hs.grad <- function(i) sum(play.ed[i,17:18])
some.coll <- function(i) sum(play.ed[i,19:21])
ba.plus <- function(i) sum(play.ed[i,22:25])

Make.ed.attain.table <- function(i) {
  this.cty <- slot(play.ed[i,1], "geography")$NAME
  data.frame(county = this.cty,
             tot = as.numeric(slot(play.ed[i,1], "estimate")),
             lt.hs = as.numeric(slot(lt.hs(i), "estimate")),
             hs.grad = as.numeric(slot(hs.grad(i), "estimate")),
             some.coll = as.numeric(slot(some.coll(i), "estimate")),
             ba.plus = as.numeric(slot(ba.plus(i), "estimate")))
}

ed.attain.list <- lapply(1:dim(play.ed)[1], Make.ed.attain.table)
ed.attain.df <- Reduce( function(x,y) full_join(x,y, all=TRUE), ed.attain.list)
```

```{r ed-attain-summary-tables-by-county}
ed.attain.df %>%
  mutate(p.lt.hs = 100*lt.hs/tot,
         p.hs.grad = 100*hs.grad/tot,
         p.some.coll = 100*some.coll/tot,
         p.ba.plus = 100*ba.plus/tot) %>%
  select(county, p.lt.hs, p.hs.grad, p.some.coll, p.ba.plus) %>%
  knitr::kable()
```

Now, calculate mean proportions of attainment across counties and use to calculate numbers out of *n*=900.

```{r proj-enroll-ed-attain}

ed.attain.df %>%
  mutate(p.lt.hs = lt.hs/tot,
         p.hs.grad = hs.grad/tot,
         p.some.coll = some.coll/tot,
         p.ba.plus = ba.plus/tot) %>%
  select(county, p.lt.hs, p.hs.grad, p.some.coll, p.ba.plus) ->
  ed.attain.by.cty

# This doesn't work because I'm still aggregating at county level
# If I were aggregating at site level it would work.
# Calculate projected enrollment by ed attain category
# ed.attain.by.cty %>%
#   mutate(proj.lt.hs = p.lt.hs*30,
#            proj.hs.grad = p.hs.grad*30,
#            proj.some.coll = p.some.coll*30,
#            proj.ba.plus = p.ba.plus*30)  %>%
#   select(county, proj.lt.hs, proj.hs.grad, proj.some.coll, proj.ba.plus) ->
#   ed.attain.by.cty

ed.attain.by.cty %>%
  knitr::kable()

# This uses mean p's so it may result in estimates > 900
ed.attain.by.cty %>%
  summarize(study.proj.lt.hs = mean(p.lt.hs),
            study.proj.hs.grad = mean(p.hs.grad),
            study.proj.some.coll = mean(p.some.coll),
            study.proj.ba.plus = mean(p.ba.plus)) %>%
  knitr::kable()
```

### Household income

It looks like table `B19001` contains "HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2013 INFLATION-ADJUSTED DOLLARS)", and there are 17 fields.

```{r household-income}
# Table B19001
hh.income <- acs.fetch(geography = play.geo, endyear = 2015, table.number = "B19001")

# If we take $25K for family of 3-4 as poverty rate
lt.25k <- function(i) sum(hh.income[i,2:5])
lt.50k  <- function(i) sum(hh.income[i,6:10])
lt.100k    <- function(i) sum(hh.income[i,11:13])
gt.100k    <- function(i) sum(hh.income[i,14:17])

Make.hh.income.table <- function(i) {
  this.cty <- slot(play.ed[i,1], "geography")$NAME
  data.frame(county = this.cty,
             tot = as.numeric(slot(hh.income[i,1], "estimate")),
             lt.25k = as.numeric(slot(lt.25k(i), "estimate")),
             lt.50k = as.numeric(slot(lt.50k(i), "estimate")),
             lt.100k = as.numeric(slot(lt.100k(i), "estimate")),
             gt.100k = as.numeric(slot(gt.100k(i), "estimate")))
}

hh.income.list <- lapply(1:dim(hh.income)[1], Make.hh.income.table)
hh.income.df <- Reduce( function(x,y) full_join(x,y, all=TRUE), hh.income.list)

hh.income.df %>%
  mutate(p.lt.25k = lt.25k/tot,
         p.lt.50k = lt.50k/tot,
         p.lt.100k = lt.100k/tot,
         p.gt.100k = gt.100k/tot) %>%
  select(county, p.lt.25k, p.lt.50k, p.lt.100k, p.gt.100k) ->
  hh.income.by.cty

hh.income.by.cty %>%
  knitr::kable()

# This uses mean p's so it may result in estimates > 900
hh.income.by.cty %>%
  summarize(study.proj.lt.25k = mean(p.lt.25k),
            study.proj.lt.50k = mean(p.lt.50k),
            study.proj.lt.100k = mean(p.lt.100k),
            study.proj.gt.100k = mean(p.gt.100k)) %>%
  knitr::kable()
```

### Percentage of Spanish speakers

It appears that table `B16007` contains info about `AGE BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 5 YEARS AND OVER`.

ACSSF	B16007	46	1	 		 	Total:	
ACSSF	B16007	46	2	 		 	5 to 17 years:	
ACSSF	B16007	46	3	 		 	Speak only English	
ACSSF	B16007	46	4	 		 	Speak Spanish	
ACSSF	B16007	46	5	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	6	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	7	 		 	Speak other languages	
ACSSF	B16007	46	8	 		 	18 to 64 years:	
ACSSF	B16007	46	9	 		 	Speak only English	
ACSSF	B16007	46	10	 		 	Speak Spanish	
ACSSF	B16007	46	11	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	12	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	13	 		 	Speak other languages	
ACSSF	B16007	46	14	 		 	65 years and over:	
ACSSF	B16007	46	15	 		 	Speak only English	
ACSSF	B16007	46	16	 		 	Speak Spanish	
ACSSF	B16007	46	17	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	18	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	19	 		 	Speak other languages	

For our purposes, I suggest we define "English speaking" as 9 + 15; "Spanish speaking" as 10+16, and "Other speaking" as sum(11:13)+sum(17:19).

```{r lang.at.home}
# Table B16007
lang.at.home <- acs.fetch(geography = play.geo, endyear = 2015, table.number = "B16007")

english <- function(i) lang.at.home[i,9] + lang.at.home[i,15]
spanish  <- function(i) lang.at.home[i,10] + lang.at.home[i,16]
other <- function(i) sum(lang.at.home[i,11:13]) + sum(lang.at.home[i,17:19])
tot <- function(i) lang.at.home[i,8] + lang.at.home[i,14]

Make.lang.at.home <- function(i) {
  this.cty <- slot(lang.at.home[i,1], "geography")$NAME
  data.frame(county = this.cty,
             tot = as.numeric(slot(tot(i), "estimate")),
             english = as.numeric(slot(english(i), "estimate")),
             spanish = as.numeric(slot(spanish(i), "estimate")),
             other = as.numeric(slot(other(i), "estimate")))
}

lang.at.home.list <- lapply(1:dim(lang.at.home)[1], Make.lang.at.home)
lang.at.home.df <- Reduce( function(x,y) full_join(x,y, all=TRUE), lang.at.home.list)

lang.at.home.df %>%
  mutate(p.english = english/tot,
         p.spanish = spanish/tot,
         p.other = other/tot) %>%
  select(county, p.english, p.spanish, p.other) ->
  lang.at.home.by.cty

lang.at.home.by.cty %>%
  knitr::kable()

# This uses mean p's so it may result in estimates != 900
lang.at.home.by.cty %>%
  summarize(study.proj.english = mean(p.english),
            study.proj.spanish = mean(p.spanish),
            study.proj.other= mean(p.other)) %>%
  knitr::kable()
```

### Better data about race/ethnicity

## Next steps

1. ~~We should confirm that our target sites collect data from the counties listed, and that the county-wide demographics are plausible~~.
2. ~~We should also see if there are *other* counties target sites collect from and consider asking them to estimate the proportion of their recruiting that comes from county A vs. county B~~. It would be fun to have a Shiny app to collect this.
3. ~~We should explore the `acs` package to grab additional demographic data, especially the indicators used in the draft grant proposal. The `choroplethr` package used in the above had demographic variables similar, but not identical to the ones we have used in the proposal~~.

## Resources

### R Session

This document was prepared in RStudio 1.0.136. Session information follows.

```{r session-info}
sessionInfo()
```
