---
title: "Set-up for site demographics"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

This document describes the preparatory steps needed to be taken prior to generating the `PLAY-recruiting-site-report.Rmd` documents.

## Install required external packages

```{r}
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
if (!require(tidycensus)) {
  install.packages("tidycensus")
}
if (!(require(ggplot2))) {
  install.packages("ggplot2")
}
if (!require(devtools)) {
  install.packages("devtools")
}
if (!(require(databraryapi))) {
  devtools::install_github("PLAY-behaviorome/databraryapi")
}
```

## Get Census API key

1. Request a Census API key using an email you have access to here: <https://api.census.gov/data/key_signup.html>
2. When you get a confirmation email, follow the link to activate the key.
3. Keep the API key handy, and run the following code in your R console:

```
tidycensus::census_api_key(key = <YOUR_CENSUS_API_KEY_SURROUNDED BY_QUOTATION_MARKS>, overwrite = TRUE, install = TRUE)
```

Substitute your Census API key (surrounded by quotations since it is a string) for `<YOUR_CENSUS_API_KEY_SURROUNDED BY_QUOTATION_MARKS>`.
This will install the API key in your local `.Renviron` environment for future use.
It will also overwrite any prior Census API keys you might have used in the past.

## Configure `databraryapi`

1. Run the following code in your console:

```
databraryapi::config_passwd()
```

This will store your Databrary password and account ID (email) in your system's secure file used for other passwords.
On Mac OS, this is called the Keychain.

2. Test that you can log on to Databrary using the API.

Run the following code in your R console:

```
databraryapi::login_db(email = <YOUR_DATABRARY_LOGIN_EMAIL>)
```

If this works, you should see a "Login successful." message, and the result of the function call, which is a Boolean value of `TRUE`.

If this fails, talk to Rick.

## Test rendering the default `PLAY-recruiting-site-report.Rmd`

1. Open `PLAY-recruiting-site-report.Rmd`.
2. Knit it by pressing the `Knit` button above the code window, or by issuing the following command in the R console:

```
rmarkdown::render("PLAY-recruiting-site-report.Rmd")
```

This should create an HTML file with information from the NYU PLAY site.

## Try rendering a set of reports

TO BE CONTINUED
