---
title: "PLAY sites demographics"
date: "`r Sys.time()`"
output: 
  html_document:
    code_folding: hide
params:
  db_login: your_account@youremail.edu
  csv_path: analysis/csv
  update_tracking_csv: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, message=FALSE, error=FALSE,
                      ccache = TRUE)

# Required packages
if(!require(tidyverse)) {
  install.packages("tidyverse")
}
if(!require(tidycensus)) {
  install.packages("tidycensus")
}

# Note: The user must acquire and install a Census API key
# See the instructions in help("census_api_key")

# Source or load dependencies
source("R/init_cap.R")

library(tidyverse) # For pipe %>% and other functions
```

# Helper functions

```{r helper-functions}
get_census_data <- function(state="PA", county="Centre", 
                            census_table="B19013_001", variable_name="median_inc",
                            get_geometry = FALSE) {
  require(tidycensus)
  require(tidyverse)
  
  if (!(is.character(state))) {
    stop("'state' must be character")
  }
  if (!(is.character(county))) {
    stop("'county' must be character")
  }
  if (!(is.character(census_table))) {
    stop("'census_table' must be character")
  }
  if (!(is.character(variable_name))) {
    stop("'variable_name' must be character")
  }
  if (stringr::str_length(state) != 2) {
    stop("'state' must be of length 2")
  }
  
  census_data <- tidycensus::get_acs(geography = "county",
                         assign(variable_name, census_table),
                         state = state,
                         county = county,
                         geometry = get_geometry,
                         cache_table = TRUE)
  # super hacky
  census_data$variable_name <- variable_name
  census_data
}
```

## Load collection site data

Create helper functions.

```{r}
get_collection_site_data <-
  function(id = "1V9RuZJNRN4lehLzRSWO0MqVclsXPw7Z-Lxq-VQOvM-A") {
    tracking_csv <-
      googledrive::drive_download(googledrive::as_id(id),
                             path = "tracking.csv",
                             overwrite = TRUE)
    tracking_df <- readr::read_csv("tracking.csv")
    tracking_df
  }

load_site_csv <- function(fn = "tracking.csv") {
  readr::read_csv(fn)
}

get_collecting_labs <- function( df = load_site_csv() ) {
  collecting_df <- dplyr::filter(df, Collection_role == "Collecting")
  collecting_df <- dplyr::select(collecting_df,
                                 Institution,
                                 `PI_Lastname`,
                                 `PI_Firstname`,
                                 `SiteCode`,
                                 City,
                                 State_prov)
  
  readr::write_csv(collecting_df, "analysis/csv/collecting-sites.csv")
  collecting_df
}

de_dupe_collecting_labs <- function(df, drop_PIs = c("Frick", "Perry")) {
  df %>%
    dplyr::filter(., !(PI_Lastname %in% drop_PIs ))
}

get_old_county_data <- function() {
  readr::read_csv("analysis/csv/city-state-county.csv")
}

get_collection_site_county_df <- function(update_tracking = FALSE) {
  if (update_tracking) {
    PLAY_all <- get_collection_site_data()
    PLAY_collecting <- get_collecting_labs(PLAY_all)
  } else {
    PLAY_collecting <- get_collecting_labs()
  }
  
  PLAY_collecting <- de_dupe_collecting_labs(PLAY_collecting)
  
  old_collecting <-
    readr::read_csv("analysis/csv/city-state-county.csv")
  dplyr::left_join(PLAY_collecting,
                   old_collecting,
                   by = c("City" = "City", "State_prov" = "State"))
}
```

Generate collection site data frame and save as CSV.

```{r}
collection_site_df <- get_collection_site_county_df()
n_counties <- 1:dim(collection_site_df)[1]
```

## White/non-white population

```{r white-non-white}
get_site_race <- function(state = "PA", county = "Centre") {
  census_table <-
    c(
      "C02003_001",
      "C02003_002",
      "C02003_003",
      "C02003_004",
      "C02003_005",
      "C02003_006",
      "C02003_007",
      "C02003_008"
    )
  variable_name <- c(
    "total",
    "one_race",
    "white",
    "black",
    "am_indian",
    "asian",
    "native_hawaiian",
    "other"
  )
  
  race_df <- get_census_data(
    state = state,
    county = county,
    census_table = census_table,
    variable_name = variable_name
  )
  
  race_df %>% 
    dplyr::select(., NAME, variable_name, estimate) %>%
    dplyr::rename(., n=estimate) %>%
    tidyr::spread(., key=variable_name, value=n)

}
```

```{r race}
make_site_race_df <- function(i, df) {
  this_site <- df[i,]
  message(paste0("Gathering race data for site: '", this_site$Institution, "'"))
  
  df_out <- get_site_race(this_site$State_prov, this_site$County)
  
  df_out$Institution <- this_site$Institution
  df_out$SiteCode <- this_site$SiteCode
  
  df_out
}
```

```{r}
collection_race_bycounty_df <- purrr::map_df(n_counties, make_site_race_df, collection_site_df)

readr::write_csv(collection_race_bycounty_df, file.path(params$csv_path, 'collection-by-county-race.csv'))
```

```{r}
tot_pop <- sum(unlist(collection_race_bycounty_df$total))
one_race <- sum(unlist(collection_race_bycounty_df$one_race))
multi_race <- tot_pop - one_race
(p_multi <- multi_race/tot_pop)

white <- sum(unlist(collection_race_bycounty_df$white)) 
black <- sum(unlist(collection_race_bycounty_df$black))
asian <- sum(unlist(collection_race_bycounty_df$asian))
other <- one_race-(white + black + asian)

(p_white <- white/tot_pop)
(p_black <- black/tot_pop)
(p_asian <- asian/tot_pop)
(p_other <- other/tot_pop)

race_df <- data.frame(race = c("multi", "white", "black", "asian", "other"),
                      p = c(p_multi, p_white, p_black, p_asian, p_other))

knitr::kable(race_df, format = 'html')
```

# Non-hispanic/Hispanic

```{r hispanic-non-hispanic}
get_hisp_non <- function(state,
                         county,
                         census_table = c("B03001_002", "B03001_003"),
                         variable_name = c("Non_hisp", "Hisp")) {

  get_census_data(state,
                  county,
                  census_table,
                  variable_name
  )
}

make_collection_site_hisp_df <- function(i, df) {
  this_site <- df[i,]
  this_institution <- 
  message(paste0("Gathering Hispanic data for site: '", this_site$Institution, "'"))
  
  df_out <- get_hisp_non(this_site$State_prov, this_site$County)
  df_out$Institution <- this_site$Institution
  df_out$SiteCode <- this_site$SiteCode
  
  df_out
}
```

```{r}
collection_hisp_bycounty_df <- purrr::map_df(n_counties, make_collection_site_hisp_df, collection_site_df)
```

```{r}
readr::write_csv(collection_hisp_bycounty_df, file.path(params$csv_path, 'collection-by-county-hispanic.csv'))
```

```{r}
collection_hisp_wide_df <- collection_hisp_bycounty_df %>%
  dplyr::filter(., variable_name %in% c("Non_hisp", "Hisp")) %>%
  tidyr::pivot_wider(., names_from = variable_name,
                     values_from = estimate)

pop_non_hisp = sum(unlist(collection_hisp_wide_df$Non_hisp))
pop_hisp = sum(unlist(collection_hisp_wide_df$Hisp))
(p_hisp = pop_hisp/(pop_hisp + pop_non_hisp))

hisp_df <- data.frame(eth = c("hisp", "non_hisp"),
                      p = c(p_hisp, 1-p_hisp))

knitr::kable(hisp_df, format = 'html')
```

The proportion of Hispanic families in the recruiting areas is `r p_hisp`.

```{r}
collection_hisp_wide_df <- collection_hisp_bycounty_df %>%
  tidyr::pivot_wider(., id_cols = c('GEOID', 'NAME', 'Institution', 'SiteCode'),
                     names_from = variable_name,
                     values_from = estimate) 

pop_non_hisp = sum(unlist(collection_hisp_wide_df$Non_hisp))
pop_hisp = sum(unlist(collection_hisp_wide_df$Hisp))
(p_hisp = pop_hisp/(pop_hisp + pop_non_hisp))

hisp_df <- data.frame(eth = c("hisp", "non_hisp"),
                      p = c(p_hisp, 1-p_hisp))

knitr::kable(hisp_df, format = 'html')
```

# Hispanic/non-Hispanic by Race

```{r hispanic-non-hispanic}
get_hisp_non_race <- function(state,
                         county,
                         census_table = c("B03002_003",
                                          "B03002_004",
                                          "B03002_005",
                                          "B03002_006",
                                          "B03002_007",
                                          "B03002_008",
                                          "B03002_009",
                                          "B03002_010",
                                          "B03002_011",
                                          "B03002_013",
                                          "B03002_014",
                                          "B03002_015",
                                          "B03002_016",
                                          "B03002_017",
                                          "B03002_018",
                                          "B03002_019",
                                          "B03002_020",
                                          "B03002_021"),
                         variable_name = c("non_hisp_white", 
                                           "non_hisp_black",
                                           "non_hisp_am_indian",
                                           "non_hisp_asian",
                                           "non_hisp_native_hawaiian",
                                           "non_hisp_other",
                                           "non_hisp_two_races",
                                           "non_hisp_two_races_incl_other",
                                           "non_hisp_three",
                                           "hisp_white",
                                           "hisp_black",
                                           "hisp_am_indian",
                                           "hisp_asian",
                                           "hisp_native_hawaiian",
                                           "hisp_other",
                                           "hisp_two_races",
                                           "hisp_two_races_incl_other",
                                           "hisp_three")) {

  get_census_data(state,
                  county,
                  census_table,
                  variable_name
  )
}

make_collection_site_hisp_race_df <- function(i, df) {
  this_site <- df[i,]
  this_institution <- 
  message(paste0("Gathering Hispanic data for site: '", this_site$Institution, "'"))
  
  df_out <- get_hisp_non_race(this_site$State_prov, this_site$County)
  df_out$Institution <- this_site$Institution
  df_out$SiteCode <- this_site$SiteCode
  
  df_out
}
```

```{r}
collection_hisp_race_bycounty_df <- purrr::map_df(n_counties, make_collection_site_hisp_race_df, collection_site_df)
```

```{r}
readr::write_csv(collection_hisp_race_bycounty_df, file.path(params$csv_path, 'collection-by-county-hispanic-race.csv'))
```

Let's try turning this into an inclusion report table.

```{r}
hisp_race_df_wider <- collection_hisp_race_bycounty_df %>%
  tidyr::pivot_wider(., id_cols = c('GEOID', 'NAME', 'Institution', 'SiteCode'),
                     names_from = variable_name,
                     values_from = estimate) 

hisp_race_df_wider <- hisp_race_df_wider %>%
  dplyr::mutate(
    .,
    non_hisp_more_than_one = non_hisp_other +
      non_hisp_two_races_incl_other +
      non_hisp_two_races +
      non_hisp_three,
    hisp_more_than_one = hisp_other +
      hisp_two_races_incl_other +
      hisp_two_races +
      hisp_three
  ) %>%
  dplyr::arrange(., SiteCode) %>%
  dplyr::select(
    .,
    -c(
      'non_hisp_other',
      'non_hisp_two_races',
      'non_hisp_two_races_incl_other',
      'non_hisp_three',
      'hisp_other',
      'hisp_two_races',
      'hisp_two_races_incl_other',
      'hisp_three'
    )
  ) %>%
  dplyr::select(
    .,
    Institution,
    SiteCode,
    NAME,
    non_hisp_white,
    non_hisp_black,
    non_hisp_am_indian,
    non_hisp_native_hawaiian,
    non_hisp_more_than_one,
    hisp_white,
    hisp_black,
    hisp_am_indian,
    hisp_native_hawaiian,
    hisp_more_than_one
  )

hisp_race_df_wider %>%
  knitr::kable(., format = 'html')

readr::write_csv(hisp_race_df_wider, file.path(params$csv_path, 'collection-by-county-hispanic-race-wide.csv'))
```

# Spanish-speakers

## Get variables

Get variables from 2019 ACS.

**NB**: The code below does not filter the ACS variables in the way I expect. The subsequent chunk selects relevant variables manually. This should be fixed in the future.

```{r get-acs19-vars}
acs_v19 <- tidycensus::load_variables(2019, "acs5", cache = TRUE)
# vars_with_spanish <- acs_v19 %>%
#   dplyr::filter(., stringr::str_detect(label, "[Ss]panish"))
# 
# census_concept <- "^LANGUAGE SPOKEN AT HOME"
# 
# lang_at_home_vars <- acs_v19 %>%
#   dplyr::filter(., stringr::str_match(concept, census_concept))
```

## Create helper functions

```{r span-speakers}
spanish_tables <-c("B16007_001", "B16007_009", "B16007_010", "B16007_011", "B16007_012", "B16007_013", "B16007_014", "B16007_015", "B16007_016", "B16007_017", "B16007_018", "B16007_019")

get_county_spanish_speakers <- function(state = "PA",
                                        county = "Centre",
                                        census_table = spanish_tables) {
  
  get_census_data(state, county, census_table, 
                  variable_name = "Spanish_speaking")
}

clean_sp_speakers <- function(df) {
  sp_clean <- df %>%
    dplyr::select(., NAME, variable, estimate) %>%
    tidyr::spread(., key = variable, value = estimate) %>%
    dplyr::group_by(NAME) %>%
    dplyr::mutate(
      .,
      English_only = B16007_009 + B16007_015,
      Spanish = B16007_010 + B16007_015,
      Other = B16007_011 + B16007_012 + B16007_013 + B16007_017 + B16007_018 + B16007_019,
      total = B16007_009 + B16007_010 +
        B16007_011 + B16007_012 +
        B16007_013 + B16007_014 +
        B16007_015 + B16007_016 +
        B16007_017 + B16007_018 +
        B16007_019,
      p_English_only = English_only / total,
      p_Spanish = Spanish / total,
      p_Other = Other / total
    ) %>%
    dplyr::select(.,
                  NAME,
                  English_only,
                  Spanish,
                  Other,
                  p_English_only,
                  p_Spanish,
                  p_Other) %>%
    tidyr::gather(., key = pop, value = n, -NAME)
  
  sp_clean
}
```

## Make county df

```{r}
make_collection_site_sp_df <- function(i, df) {
  this_site <- df[i,]
  message(paste0("Gathering Spanish speaker data for site: '", this_site$Institution, "'"))
  
  this_sp_df <- get_county_spanish_speakers(this_site$State_prov, this_site$County)
  
  this_sp_df$Institution <- this_site$Institution
  this_sp_df$SiteCode <- this_site$SiteCode
  
  clean_sp_speakers(this_sp_df)
}
```

Gather data by county.

```{r}
collection_sp_bycounty_df <- purrr::map_df(n_counties, make_collection_site_sp_df, collection_site_df)
```

Save data.

```{r}
readr::write_csv(collection_sp_bycounty_df, file.path(params$csv_path, 'collection-by-county-spanish.csv'))
```

Calculate summary stats across data collection sites.

```{r}
collection_sp_wide_df <- collection_sp_bycounty_df %>%
  dplyr::filter(., pop %in% c('English_only', 'Spanish', 'Other')) %>%
  tidyr::pivot_wider(., names_from = pop,
                     values_from = n)

pop_eng <- sum(unlist(collection_sp_wide_df$English_only))
pop_span <- sum(unlist(collection_sp_wide_df$Spanish))
pop_oth <- sum(unlist(collection_sp_wide_df$Other))
pop_total <- pop_eng + pop_span + pop_oth
p_span <- pop_span/pop_total

language_df <- tibble::tibble(pop_eng, pop_span, pop_oth, pop_total, p_span)

language_df %>%
  knitr::kable(., format = 'html')
```

The proportion of Spanish-speaking families in the recruitment areas is `r p_span`.

# Educational attainment

```{r ed-attain-functions}
get_ed_attain <- function(state="PA", county="Centre") {
  census_table <- c("B15003_001", # total
                    "B15003_002", # none
                    "B15003_003", # <K
                    "B15003_004", # K
                    "B15003_005", # 1st
                    "B15003_006", # 2nd
                    "B15003_007", # 3rd
                    "B15003_008", # 4th
                    "B15003_009", # 5th
                    "B15003_010", # 6th
                    "B15003_011", # 7th
                    "B15003_012", # 8th
                    "B15003_013", # 9th
                    "B15003_014", # 10th
                    "B15003_015", # 11th
                    "B15003_016", # 12th
                    "B15003_017", # HS
                    "B15003_018", # GED
                    "B15003_019", # Coll<1
                    "B15003_020", # Coll>1
                    "B15003_021", # AA
                    "B15003_022", # BA
                    "B15003_023", # MA
                    "B15003_024", # Prof
                    "B15003_025")  # PhD
  variable_name <- "ed_attain"
  get_census_data(state = state, county = county, 
                  census_table = census_table,
                  variable_name = variable_name)
}
```

```{r}
make_ed_attain_wide <- function(df) {
  d0 <- dplyr::select(df, NAME, variable, estimate)
  d1 <- tidyr::spread(d0, key = variable, value = estimate)
  d2 <- dplyr::rename(d1, total = B15003_001, # total
                          none = B15003_002,  # none
                          preK = B15003_003,
                      K = B15003_004,
                      g01 = B15003_005, # 1st
                      g02 = B15003_006, # 2nd
                      g03 = B15003_007, # 3rd
                      g04 = B15003_008, # 4th
                      g05 = B15003_009, # 5th
                      g06 = B15003_010, # 6th
                      g07 = B15003_011, # 7th
                      g08 = B15003_012, # 8th
                      g09 = B15003_013, # 9th
                      g10 = B15003_014, # 10th
                      g11 = B15003_015, # 11th
                      g12 = B15003_016, # 12th
                      hs  = B15003_017, # HS
                      ged = B15003_018, # GED
                      coll_lt_1 = B15003_019, # Coll<1
                      coll_gt_1 = B15003_020, # Coll>1
                      aa = B15003_021, # AA
                      ba = B15003_022, # BA
                      ma = B15003_023, # MA
                      prof = B15003_024, # Prof
                      phd = B15003_025)  # PhD
  d2
}

simplify_ed_attain <- function(df) {
  # Calculate aggregate variables
  df_new <- df
  df_new <- dplyr::mutate(df_new,
                          lt_hs = none + preK + g01 + g02 +
                            g03 + g04 + g05 + g06 + g07 +
                            g08 + g09 + g10 + g11 + g12,
                          hs_grad = hs + ged,
                          some_coll = coll_lt_1 + coll_gt_1 + aa,
                          ba_plus = ma + prof + phd)
  df_new <- dplyr::select(df_new,
                          NAME, total, lt_hs, hs_grad, some_coll, ba,
                          ba_plus)
  df_new
}

make_site_ed <- function(i, df) {
  this_site <- df[i,]
  message(paste0("Gathering educational attainment data for site: '", this_site$Institution, "'"))
   
  this_ed_df <- get_ed_attain(this_site$State_prov, this_site$County)
  
  this_ed_wide_df <- make_ed_attain_wide(this_ed_df)
  
  simplify_ed_attain(this_ed_wide_df)
}
```

Gather data by county.

```{r}
collection_ed_bycounty_df <- purrr::map_df(n_counties, make_site_ed, collection_site_df)
```

Save data.

```{r}
readr::write_csv(collection_ed_bycounty_df, file.path(params$csv_path, 'collection-by-county-education.csv'))
```

Calculate summary variables.

```{r}
tot_ed <- sum(unlist(collection_ed_bycounty_df$total))
lt_hs <- sum(unlist(collection_ed_bycounty_df$lt_hs))
hs_grad <- sum(unlist(collection_ed_bycounty_df$hs_grad))
some_coll <- sum(unlist(collection_ed_bycounty_df$some_coll))       

ba <- sum(unlist(collection_ed_bycounty_df$ba))
ba_plus <- sum(unlist(collection_ed_bycounty_df$ba_plus))

(p_lt_hs <- lt_hs/tot_ed)
(p_hs_grad <- hs_grad/tot_ed)
(p_some_coll <- some_coll/tot_ed)
#(p_ba <- ba/tot_ed)
#(p_ba_plus <- ba_plus/tot_ed)
(p_ba_ba_plus <- (ba+ba_plus)/tot_ed)

ed_df <- data.frame(ed = c("lt_hs", "hs_grad", "some_coll", "ba_plus"),
                     p = c(p_lt_hs, p_hs_grad, p_some_coll, p_ba_ba_plus))

knitr::kable(ed_df, 'html')
```

# Household income

Find tables.

```{r}
acs_v19 %>% 
  filter(., str_detect(concept, "HOUSEHOLD INCOME")) %>% 
  head(17)
```

Looks like tables B19001_001:B19001_017

```{r household-income}
get_household_inc <- function(state = "PA", county = "Centre") {
  census_table <- c(
    "B19001_001",
    # total
    "B19001_002",
    # none
    "B19001_003",
    # <K
    "B19001_004",
    # K
    "B19001_005",
    # 1st
    "B19001_006",
    # 2nd
    "B19001_007",
    # 3rd
    "B19001_008",
    # 4th
    "B19001_009",
    # 5th
    "B19001_010",
    # 6th
    "B19001_011",
    # 7th
    "B19001_012",
    # 8th
    "B19001_013",
    # 9th
    "B19001_014",
    # 10th
    "B19001_015",
    # 11th
    "B19001_016",
    # 12th
    "B19001_017"
  )  # PhD
  
  variable_name <- c(
    "total",
    "<10K", "10-15K",
    "15-20K",
    "20-25K",
    "25-30K",
    "30-35K",
    "35-40K",
    "40-45K",
    "45-50K",
    "50-60K",
    "60-75K",
    "75-100K",
    "100-125K",
    "125-150K",
    "150-200K",
    ">200K"
  )
  
  get_census_data(
    state = state,
    county = county,
    census_table = census_table,
    variable_name = variable_name
  )
}

simplify_inc <- function(df) {
  # Calculate aggregate variables
  df_new <- df
  df_new <- dplyr::mutate(df_new,
                          lt_25K = `<10K` + `10-15K` + `15-20K` + `20-25K`,
                          b25_50K = `25-30K` + `30-35K` + `35-40K` + `40-45K` + `45-50K`,
                          b50_100K = `50-60K` + `60-75K` + `75-100K`,
                          gt_100K = `100-125K` + `125-150K` + `150-200K` + `>200K`)
  df_new <- dplyr::select(df_new,
                          NAME, total, lt_25K, b25_50K, b50_100K, gt_100K)
  df_new
}
```

```{r}
make_site_inc <- function(i, df) {
  this_site <- df[i, ]
  message(paste0("Gathering household income data for site: '", this_site$Institution, "'"))
  
  this_inc_df <-
    get_household_inc(this_site$State_prov, this_site$County) %>%
    dplyr::select(., -variable, -GEOID, -moe) %>%
    dplyr::rename(., n = estimate)
  
  this_inc_wide <- tidyr::pivot_wider(this_inc_df, names_from = variable_name,
                                      values_from = n)
  
  this_inc_wide %>% simplify_inc(.)
}
```

```{r}
collection_inc_bycounty_df <- purrr::map_df(n_counties, make_site_inc, collection_site_df)
```

Save data.

```{r}
readr::write_csv(collection_inc_bycounty_df, file.path(params$csv_path, 'collection-by-county-income.csv'))
```

```{r}
tot_inc <- sum(unlist(collection_inc_bycounty_df$total))
lt_25K <- sum(unlist(collection_inc_bycounty_df$lt_25K))
b25_50K <- sum(unlist(collection_inc_bycounty_df$b25_50K))
b50_100K <- sum(unlist(collection_inc_bycounty_df$b50_100K))
gt_100K <- sum(unlist(collection_inc_bycounty_df$gt_100K))

(p_lt_25K <- lt_25K/tot_inc)
(p_b25_50K <- b25_50K/tot_inc)
(p_b50_100K <- b50_100K/tot_inc)
(p_gt_100K <- (gt_100K)/tot_inc)

inc_df <- data.frame(income = c("lt_25K", "b25-50K", "b50-100K", "gt_100K"),
                     p = c(p_lt_25K, p_b25_50K, p_b50_100K, p_gt_100K))

knitr::kable(inc_df, 'html')
```



