---
title: "spanish_speaking"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  pdf_document:
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path = "img/",
                      fig.align = "center")
library(tidyverse)
library(choroplethr)
library(choroplethrMaps)

csv.dir <- 'analysis/csv/'
R.dir <- 'R/'

source(paste0(R.dir, "Cap_all.R"))

# ggplot
play.palette <- scale_fill_manual(values=c("blue2", "firebrick2", "chartreuse2", "darkorchid2", "yellow", "gray"))

play.theme <-   
  theme_classic() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = rel(1.2)),
        axis.title = element_text(size = rel(1.5), face ="bold"),
        axis.text.x = element_text(size = rel(1.2)),
        axis.text.y = element_text(size = rel(.7)),
        axis.line = element_blank(),
        axis.title.y = element_blank(),
#        plot.margin=unit(c(1,1,1,1),"cm"))
        plot.margin=unit(c(0,.5,0,0),"cm"))
```

## Spanish speakers

```{r download-demographic-data}
counties <- read.csv(paste0(csv.dir, "city-state-county.csv"), 
                     stringsAsFactors = FALSE)

data("county.regions")
counties <- left_join(counties, county.regions)
demog <- get_county_demographics(endyear=2013, span=5)
county.demo <- left_join(counties, demog)

# Recapitalize county
county.demo$County <- unlist(lapply(county.demo$County, Cap_all))
# Hack District Of columbia...TODO(ROG): Fix Cap_all()
county.demo$County[county.demo$County == "District Of columbia"] = "District of Columbia"
```
## Spanish speaking

To gather additional ACS data, we must create a specific ACS 'geometry'.

```{r generate-site-geography}
state.fips <- as.numeric(county.demo$state.fips.character)
county.fips <- as.numeric(substr(county.demo$county.fips.character,3,5))
play.geo <- geo.make(state = state.fips, county = county.fips)

Make.county.geo <- function(i, df) {
  geo.make(state = as.numeric(df$state.fips.character[i]),
              county =
             as.numeric(substr(county.demo$county.fips.character[i],4, 6)))
}

cty <- 1
# Generate name for county-level geography
geo.name <- paste0(county.demo$Site.code[cty], "_",
                   county.demo$county.name[cty], "_", county.demo$State[cty])

# Create geography and assign to generated name
assign(geo.name, Make.county.geo(cty, county.demo))
```


It appears that ACS table `B16007` contains info about `AGE BY LANGUAGE SPOKEN AT HOME FOR THE POPULATION 5 YEARS AND OVER`.

ACSSF	B16007	46	1	 		 	Total:	
ACSSF	B16007	46	2	 		 	5 to 17 years:	
ACSSF	B16007	46	3	 		 	Speak only English	
ACSSF	B16007	46	4	 		 	Speak Spanish	
ACSSF	B16007	46	5	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	6	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	7	 		 	Speak other languages	
ACSSF	B16007	46	8	 		 	18 to 64 years:	
ACSSF	B16007	46	9	 		 	Speak only English	
ACSSF	B16007	46	10	 		 	Speak Spanish	
ACSSF	B16007	46	11	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	12	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	13	 		 	Speak other languages	
ACSSF	B16007	46	14	 		 	65 years and over:	
ACSSF	B16007	46	15	 		 	Speak only English	
ACSSF	B16007	46	16	 		 	Speak Spanish	
ACSSF	B16007	46	17	 		 	Speak other Indo-European languages	
ACSSF	B16007	46	18	 		 	Speak Asian and Pacific Island languages	
ACSSF	B16007	46	19	 		 	Speak other languages	

For our purposes, I suggest we define "English speaking" as 9 + 15; "Spanish speaking" as 10+16, and "Other speaking" as sum(11:13)+sum(17:19).

```{r lang.at.home}
# Table B16007
lang.at.home <- acs.fetch(geography = play.geo, 
                          endyear = 2015, table.number = "B16007")

english <- function(i) lang.at.home[i,9] + lang.at.home[i,15]
spanish  <- function(i) lang.at.home[i,10] + lang.at.home[i,16]
other <- function(i) sum(lang.at.home[i,11:13]) + sum(lang.at.home[i,17:19])
tot <- function(i) lang.at.home[i,8] + lang.at.home[i,14]

Make.lang.at.home <- function(i) {
  this.cty <- slot(lang.at.home[i,1], "geography")$NAME
  data.frame(county = this.cty,
             tot = as.numeric(slot(tot(i), "estimate")),
             english = as.numeric(slot(english(i), "estimate")),
             spanish = as.numeric(slot(spanish(i), "estimate")),
             other = as.numeric(slot(other(i), "estimate")))
}

lang.at.home.list <- lapply(1:dim(lang.at.home)[1], Make.lang.at.home)
lang.at.home.df <- Reduce(function(x,y) full_join(x,y, all=TRUE),
                          lang.at.home.list)

lang.at.home.df %>%
  mutate(`English only` = english/tot*100,
         Spanish = spanish/tot*100,
         Other = other/tot*100) %>%
  select(county, `English only`, Spanish, Other) ->
  lang.at.home.by.cty
```

```{r spanish-speaking-table}
lang.at.home.by.cty %>%
  knitr::kable()
```

```{r spanish-speaking-plot}
lang.at.home.by.cty %>%
  gather(key = language, value = proportion.pop, -county) %>%
  mutate(language = factor(language, levels = c("Other", "Spanish", "English only"))) ->
  lang.at.home.by.cty.gathered

lang.at.home.by.cty %>%
  select(county, `English only`) ->
  lang.sortlist

left_join(lang.at.home.by.cty.gathered, lang.sortlist) %>%
  arrange(`English only`) %>%
  mutate(county = factor(county, unique(county))) %>%
  ggplot() +
  aes(x = county, y = proportion.pop, fill = language) +
  geom_col() +
  coord_flip() +
  play.theme +
  play.palette +
  scale_y_continuous(expand=c(0,0)) +
  ylab("Percent of population")
```
