---
title: "Demographics with `tidycensus`"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output:
  html_document:
    code_folding: hide
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
params: 
  census_api_key: "abcdefg"
---


# Purpose

This document describes Rick Gilmore's explorations of the `tidycensus` package.
The goal is to determine whether this package can meet some of the needs for demographic visualization and analysis as outlined <demographics-2.0.html>.
The package site on GitHub is here: <https://github.com/walkerke/tidycensus>.
This exploration relies heavily on the documentation provided here: <https://walkerke.github.io/tidycensus/articles/basic-usage.html>.

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(tidycensus)) {
  install.packages('tidycensus')
}
if (!require(tidyverse)) {
  install.packages('tidyverse')
}
if (!require(ggplot2)) {
  install.packages('ggplot2')
}
if (!require(viridis)) {
  install.packages('viridis')
}
```

## Load `tidycensus` and associated packages

This document installs `tidycensus`, `tidyverse`, `ggplot2`, and `viridis` if these are not already installed when the document is knit.

```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(viridis)
```

## Add Census API key

According to the `tidycensus` documentation, the following command writes the Census API key to the `.Renviron` environment variable. for use in future sessions if `params$census_api_key` is defined in the `rmarkdown::render()` call to the document.
For example, run `rmarkdown::render("tidycensus.Rmd", params=list(census_api_key=<YOUR_API_KEY>))` substituting your API key for \<YOUR_API_KEY\>.

```{r store-api-key, eval=FALSE}
census_api_key(key = params$census_api_key, install = TRUE)
```

For future runs, execute the following:

```{r read-api-key, eval=TRUE}
readRenviron("~/.Renviron")
```

# Testing functionality

## Focus on PA

```{r}
pa <- get_acs(geography = "county", 
              variables = c(medincome = "B19013_001"), 
              state = "PA")

pa
```

Wow, super.
Let's see if we can generate a map.

```{r}
pa_county_income <- get_acs(geography = "county", 
                            c(medincome = "B19013_001"), state = "PA", 
                            geometry = TRUE)

ggplot(pa_county_income) + 
  geom_sf(aes(fill = estimate), color = NA) + 
  coord_sf(datum = NA) + 
  theme_minimal() + 
  scale_fill_viridis_c()
```

Wow again.
I wonder if I can filter and show just the counties around Centre County.

## Map of Centre County plus surrounding

```{r}
centre_plus <- "Centre|Clearfield|Clinton|Union|Mifflin|Huntingdon|Blair"

centre_county_plus_income <- pa_county_income %>%
  dplyr::filter(stringr::str_detect(NAME, centre_plus))

centre_county_plus_income %>%
  ggplot(.) + 
  geom_sf(aes(fill = estimate), color = NA) + 
  coord_sf(datum = NA) + 
  theme_minimal() + 
  scale_fill_viridis_c()  
```

Sweet.
While we're here, let's do a local version of the plot with error estimates.

## Plot of Centre County plus surrounding median incomes

```{r}
centre_county_plus_income %>%
  mutate(NAME = gsub(" County, Pennsylvania", "", NAME)) %>%
  ggplot(aes(x = estimate, y = reorder(NAME, estimate))) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point(color = "red", size = 3) +
  labs(title = "Household income by county in Central Pennsylvania",
       subtitle = "2013-2017 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)")
```

