---
title: "PLAY Planned Enrollment"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output: 
  github_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      error = FALSE,
                      message = FALSE,
                      fig.path = "img/")

# Install packman for subsequent package management
if (!("pacman" %in% installed.packages()[,])) {
    install.packages("pacman")
}

# Load packages
pacman::p_load(ggplot2, dplyr, stringr, choroplethr, choroplethrMaps)

# Paths
csv.dir <- "analysis/csv/"

# Helper functions
source("analysis/R/Cap.all.R")
```

## Background/Rationale

This notebook describes the projected enrollment across the $n$=30 data collection sites based upon Census data about the race and ethnicity composition of the surrounding counties.

## Generate data files

Load county names, and add county FIPS from `choroplethr` package's `county.regions` dataset. With this, I can extract demographic data using the `get_county_demographics()` function.

Note, that to access Census API, one must first acquire an API key from <http://www.census.gov/developers/>. Then, one must run `api.key.install()` to avoid throwing the API error; alternatively give the api key value as `key = <api.key>`.

```{r generate-data-files}
counties <- read.csv(paste0(csv.dir, "city-state-county.csv"), stringsAsFactors = FALSE)

# convert counties so can pull FIPS codes
#counties$County <- tolower(counties$County)

# Load county data from choroplethr
# Could also use acs package to get updated info.
data("county.regions")

counties <- left_join(counties, county.regions)

demo <- get_county_demographics(endyear=2013, span=5)

county.demo <- left_join(counties, demo)

# Recapitalize county
county.demo$County <- unlist(lapply(county.demo$County, Cap.all))
```

## Tabular summary

There are `r dim(county.demo)[1]` counties among the `r length(unique(county.demo$Site.code))` data collection sites. Here are their demographics.

```{r tabular-summary}
# Select collecting sites only
county.demo %>%
  filter(Collecting == "Collecting") ->
  county.demo

county.demo %>%
  arrange(US.Region, Site.code, State, County) %>%
  select(US.Region, Site.code, State, County, total_population, percent_white,
         percent_black, percent_asian, percent_hispanic) %>%
  knitr::kable()
```

## Calculating planned enrollment by site

We calculate expected proportions of participants based on county-level race percentages, assuming an $n$=30. Take the floor of `N_white` and the ceiling of others since we hope to increase sample diversity.

```{r planned-enrollment-by-race}
county.demo %>%
  arrange(US.Region, Site.code, State, County) %>%
  select(US.Region, Site.code, State, County, total_population,
         percent_white, percent_black, percent_asian,
         percent_hispanic, multi) -> 
  county.race.ethnicity

county.race.ethnicity %>%
  mutate(N_white = floor(percent_white*30/100),
         N_black = ceiling(percent_black*30/100),
         N_asian = ceiling(percent_asian*30/100),
         N_hispanic = ceiling(percent_hispanic*30/100)) %>%
  select(US.Region, Site.code, State, County, N_white, N_black, N_asian, N_hispanic, multi) %>%
  mutate(N_site = N_white + N_black + N_asian + N_hispanic) ->
  county.planned.enrollment
  
county.planned.enrollment %>%
  knitr::kable()
```

Separate out sites that recruit from multiple counties.

```{r planned-enrollment-across-sample}
county.planned.enrollment %>%
  filter(multi == "yes") ->
  county.demo.multi

county.planned.enrollment %>%
  filter(multi != "yes") ->
  county.demo.single
```

There are `r length(unique(county.demo.multi$Site.code))` sites that recruit from more than one county or metropolitan statistical area. For reasons I can't quite determine, the following code fails to create the expected site-by-site summaries.

```{r compute-multisite-projections, eval=FALSE}
county.demo.multi %>%
  group_by(Site.code) %>%
  summarize(N_white_proj = mean(N_white),
            N_black_proj = mean(N_black),
            N_asian_proj = mean(N_asian),
            N_hisp_proj  = mean(N_hispanic)) %>%
  knitr::kable()
```

So, I compute them individually. Take optimistic projections for the non-white sample.

```{r compute-multisite-by-site}
sites <- unique(county.planned.enrollment$Site.code)

# Compute average of all counties at a site
Project.site.demo <- function(this.site, df) {
  df %>%
  filter(Site.code == this.site) %>%
  summarize(N_black = ceiling(mean(N_black)),
            N_asian = ceiling(mean(N_asian)),
            N_hisp  = ceiling(mean(N_hispanic)),
            N_white = 30 - N_black - N_asian - N_hisp,
            Pct_white = 100*N_white/(N_white + N_black + N_asian + N_hisp),
            Pct_black = 100*N_black/(N_white + N_black + N_asian + N_hisp),
            Pct_asian = 100*N_asian/(N_white + N_black + N_asian + N_hisp),
            Pct_hisp = 100*N_hisp/(N_white + N_black + N_asian + N_hisp)) ->
    df2
  #df2$US.Region = unique(df$US.Region)
  df2$Site.code = this.site
  #df2$State = unique(df$State)
  df2
}

# Apply the function across the list of sites
demo.proj.list <- lapply(sites, Project.site.demo,
                         df=county.planned.enrollment)
demo.proj <- Reduce(function(x,y) merge(x,y, all = TRUE), x = demo.proj.list)

demo.proj %>%
  select(Site.code, Pct_white, Pct_black, Pct_asian, Pct_hisp) %>%
  knitr::kable()

demo.proj %>%
  summarize(Tot_white = sum(N_white),
            Tot_black = sum(N_black),
            Tot_asian = sum(N_asian),
            Tot_hisp  = sum(N_hisp),
            Tot_all   = Tot_white + 
              Tot_black + Tot_asian + Tot_hisp) ->
  demo.proj.total

demo.proj.total %>%
  knitr::kable()

demo.proj.total %>%
  summarize(Pct_white = 100*Tot_white/Tot_all,
            Pct_black = 100*Tot_black/Tot_all,
            Pct_asian = 100*Tot_asian/Tot_all,
            Pct_hisp  = 100*Tot_hisp/Tot_all) %>%
  knitr::kable()
```

| Race/Eth | Min Pct | Max Pct | Mean Pct | Median Pct |
|----------|---------|---------|----------|------------|
| White    | `r min(demo.proj$Pct_white)` |`r max(demo.proj$Pct_white)`| `r mean(demo.proj$Pct_white)`|`r median(demo.proj$Pct_white)`|
| Black    | `r min(demo.proj$Pct_black)` |`r max(demo.proj$Pct_black)`| `r mean(demo.proj$Pct_black)`|`r median(demo.proj$Pct_black)`|
| Asian    | `r min(demo.proj$Pct_asian)` |`r max(demo.proj$Pct_asian)`| `r mean(demo.proj$Pct_asian)`|`r median(demo.proj$Pct_asian)`|
| Hispanic    | `r min(demo.proj$Pct_hisp)` |`r max(demo.proj$Pct_hisp)`| `r mean(demo.proj$Pct_hisp)`|`r median(demo.proj$Pct_hisp)`|

## Resources

This document was prepared in RStudio 1.0.143. Session information follows.

```{r session-info}
sessionInfo()
```
