---
title: "play-launch-group"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output:
  github_document:
  html_document:
    toc: true
    toc_levels: 2
    toc_floating: true
    code_folding: hide
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      error = FALSE,
                      message = FALSE,
                      fig.path = "img/")

# Install packman for subsequent package management
if (!("pacman" %in% installed.packages()[,])) {
    install.packages("pacman")
}

# Load packages
pacman::p_load(ggplot2, dplyr, stringr)

# Paths
csv.dir <- "analysis/csv/"

# Helper functions
source("analysis/R/Cap.all.R")
```

## Background

This report describes the characteristics of the data collection and coding sites and the launch group members.

## Generate data file

Open the project roster.

```{r open-roster-file}
launch_group <- read.csv(paste0(csv.dir, "play-roster.csv"), stringsAsFactors = FALSE)
```

## Summarize characteristics of launch group members

To determine R15 status, I compared list of institutions that are *ineligible* according to this list from NIH:

<https://grants.nih.gov/grants/funding/R15_eligibility_2016-2017.pdf>

I edited the play-roster.csv file accordingly.

```{r launch-group-member-summaries}
# New/Not_new
launch_group %>% 
  group_by(New) %>% 
  summarise(N = n()) %>%
  knitr::kable()

# Gender
launch_group %>% 
  group_by(Gender) %>% 
  summarise(N = n()) %>%
  knitr::kable()

# White/non
launch_group %>% 
  group_by(Race_eth) %>% 
  summarise(N = n()) %>%
  knitr::kable()

# Institution Type
launch_group %>% 
  group_by(Institution_Type) %>% 
  summarise(N = n()) %>%
  knitr::kable()

# R-15 eligible
launch_group %>% 
  group_by(R15_eligible) %>% 
  summarise(N = n()) %>%
  knitr::kable()

```

